---
output:
  knitrBootstrap::bootstrap_document:
    title: "Test file"
    theme: simplex
    highlight: "Brown Paper"
    theme.chooser: FALSE
    highlight.chooser: FALSE
    menu: FALSE
---


```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
source("R/1_global.R")         
source("R/2_functions.R")         
source("R/3_load-and-clean.R")
source("R/4_figures.R")          
source("R/5_tables.R")
```


## Plugin code
Box 1.  
```{r,eval=FALSE}
dLength<- function(x, n, k, linf,length1,dT,live)
    {# FABENS MODEL WITH MODFICATION	
    # x    number of replicates, column index used by sapply 
    # n    number in super population
    # k    vector of growth coefficients for individual fish
    # linf    vector of length at infinity for individual fish
    # length1    vector of length at t-dt for individual fish
    # dT    change in time
    # live    is the fish alive or dead?

    length2<-(linf[,x]-length1[,x]*(1-exp(-k[,x]*dT))+ 
    length1[,x]) * live[,x]

	return(length2)# return the predicted length
	}
```


Box 2.  Caption
```{r,eval=FALSE}
dWeight<- function(x,n,len,a_prime,b,er,live)
	{
    # x       number of replicates, column index used by sapply 
	# n       number in super population
	# len     vector of current lengths
	# a_prime intercept of length weight equation
	# b       allometric scaling coefficient
 	# er      cv; log normally distributed uncertainty
	# live    is the fish alive or dead?

	weight<-exp(rnorm(n, a_prime+b*len[,x],er))*live[,x]

    return(weight)# return the predicted weight
	}
```



## Figures

figures(1) # study area
savePlot("./figures/figure-01.png",type="png")

figures(2) # length weight plot for upper and lower basin
savePlot("./figures/figure-02.png",type="png")

figures(3) # Change in length (y-axis) versus time at large (x-axis) plot for upper and lower basin
savePlot("./figures/figure-03.png",type="png")

figures(5)# PLOT TIME AND GROWTH
savePlot("./figures/figure-05.png",type="png")

## Tables

tbl1<- tables("tbl1") # length weight relationship by basin
write.csv(tbl1, "./tables/table-01.csv")
	
tbl2<- tables("tbl2") # growth model selection
write.csv(tbl2, "./tables/table-02.csv")

tbl3<- tables("tbl3") # growth model parm estimates
write.csv(tbl3, "./tables/table-03.csv")	
	
tbl5<- tables("tbl5")# growth model parm estimates-correlated linf and k
write.csv(tbl5, "./tables/table-05.csv")		


























 

